---
title: "Data Preprocessing"
author: "Ivan Timothy Halim"
date: "12/3/2020"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(lubridate)
```

# Data Preprocessing

## Import Data

Data is acquired from New York Stock Exchange on Kaggle ( https://www.kaggle.com/dgawlik/nyse ). We will only use data from January to March of 2015 for illustration.

* `date`: date
* `symbol`: symbol of company stock
* `open`: price at the open of the day
* `close`: price at the end of the day
* `low`: lowest price of the day
* `high`: highest price of the day
* `volume`: number of transaction at the day

```{r}
nyse <- read_csv(here("data", "prices.csv"))

nyse <- nyse %>%
    mutate(date = ymd(date)) %>%
    filter(year(date) == 2015,
           month(date) %in% c(1:3))
head(nyse)
```

To get clearer name of company, let's import the Ticker Symbol and Security.

```{r}
securities <- read_csv(here("Data", "securities.csv"))
securities <- securities %>%
    select(`Ticker symbol`, Security) %>%
    rename(stock = `Ticker symbol`)
head(securities)
```

Let's say I have assets in 3 different stocks. I will randomly choose the stocks.

```{r}
set.seed(13)
selected_stock <- sample(nyse$symbol, 3)

nyse <- nyse %>%
    filter(symbol %in% selected_stock)
head(nyse)
```

## Calculate Returns

Let's calculate the daily returns

```{r}
nyse <- nyse %>%
    select(date, symbol, close) %>%
    group_by(symbol) %>%
    rename(price = close) %>%
    mutate(price_prev = lag(price),
           returns = (price - price_prev)/price_prev) %>%
    slice(-1) %>%
    ungroup()

head(nyse)
```

Let's calculate the mean return of each stock

```{r}
mean_stock <- nyse %>%
    group_by(symbol) %>%
    summarise(mean = mean(returns))
head(mean_stock)
```

The value of ![equation](https://latex.codecogs.com/gif.latex?R_f) is acquired from the latest interest rate on a three-month U.S. Treasury bill. Since the data is from 2016, we will use data from 2015 (Use data from March 27, 2015), which is 0.04%. The rate is acquired from https://ycharts.com/indicators/3_month_t_bill.

```{r}
rf <- 0.04/100
```

## Covariance Matrix Between Portofolio

Calculate the covariance matrix between portofolio. First, we need to separate the return of each portofolio into several column by spreading them.

```{r}
nyse_wide <- nyse %>%
    pivot_wider(id_cols = c(date, symbol), names_from = symbol, values_from = returns) %>%
    select(-date)

# Create Excess Return
for (symbol in unique(nyse$symbol)) {
    nyse_wide[symbol] <- nyse_wide[symbol] - as.numeric(mean_stock[mean_stock$symbol == symbol, "mean"])
}

head(nyse_wide)
```

Create the covariance matrix

```{r}
(nyse_cov <- cov(x = nyse_wide))
```

Let's save our data for future simulations

```{r}
saveRDS(nyse, here("results", "nyse_prices.rds"))
saveRDS(securities, here("results", "nyse_securities.rds"))
saveRDS(mean_stock, here("results", "nyse_mean_returns.rds"))
saveRDS(rf, here("results", "risk_free_rate.rds"))
saveRDS(nyse_cov, here("results", "nyse_covariance.rds"))
```